Optimal updating factor in Wang-Landau and metadynamics simulations

### 2016-04-12 doc/invt.tex
invt.tex: editing conclusion

Going back to the problem of
minimizing Eq. \eqref{eq:error_sum}
for the single-bin scheme.
%
Note that Eq. \eqref{eq:dxdt_WL}
is of the form of Eq. \eqref{eq:dxdt_alpha}.
%
Thus, with $\lambda = 1$,
Eq. \eqref{eq:alpha_opt}
will optimize each term $\langle x_i^2 \rangle$,
and hence the sum.
%
In other words, the optimal $\alpha(t)$
is given by
\begin{equation}
\alpha(t)
=
\frac{    1     }
     {  t + t_0 }
,
\label{eq:alpha_invt1}
\end{equation}
and Eq. \eqref{eq:alpha_invt}
is the case of $t_0 = 0$.

@@

We now consider the multiple-bin scheme,
Eq. \eqref{eq:mbin_update}.
%
In this case,
upon a visit to bin $j$,
we update not only the bias potential at bin $j$,
but also those at a few neighboring bins, $i$'s.
%
The update magnitudes are given by a matrix
$\mathbf w$
with elements $w_{ij}$.
%
Then, Eq. \eqref{eq:vt_diffeq} is generalized to
\begin{equation}
  \dot v_i(t)
  =
  \alpha(t) \sum_{j=1}^n \, w_{ij} \frac{ h_j(t) } { p_j }.
  \label{eq:vt_diffeq_mbin}
\end{equation}

@@

The matrix $\mathbf w$ is not arbitrary.
%
A necessary condition to sample
the desired distribution
$\mathbf p = (p_1, \dots, p_n)$
is the following.
%
In Eq. \eqref{eq:vt_diffeq_mbin},
if $h_j(t)$ were the same as $p_j$,
the rates of change $\dot v_i(t)$'s
should be independent of $i$.
%
This allows $v_i(t)$'s to grow uniformly
in the asymptotic regime.
%
Thus, by a proper scaling of $\alpha(t)$,
we may write this condition as
%
\begin{equation}
  \sum_{j = 1}^n w_{ij} = 1.
  \label{eq:w_sumj}
\end{equation}
%
The above condition means that $(1, \dots, 1)^T$
is a right eigenvector of $\mathbf w$
with eigenvalue $1$.
%
Thus, the transpose $\mathbf w^T$
resembles a transition matrix,
although the matrix elements can be negative.

@@

Further, by Eq. \eqref{eq:w_detailedbalance},
we can define a symmetric matrix $\hat{\mathbf w}$
as $\hat w_{ij} = \sqrt{p_i/p_j} \, w_{ij}$
that can be diagonalized\cite{vankampen}
with a set of orthonormal eigenvectors:
%
$\sum_{i = 1}^n \varphi_{ki} \, \hat w_{ij} = \lambda_k \, \varphi_{kj}$.
%
Thus,
in terms of diagonalizing $\mathbf w$, we have
\begin{equation}
  \sum_{i = 1}^n \phi_{ki} \, w_{ij} = \lambda_k \, \phi_{kj},
  \label{eq:eig_w}
\end{equation}
where the eigenvectors
$$\phi_{ij} \equiv \sqrt{p_j} \, \varphi_{ij}$$
satisfy the orthonormal conditions\cite{vankampen}:
%
\begin{align}
\sum_{k = 1}^n \phi_{ki} \, \phi_{kj}
&= \delta_{ij} \, p_i,
\label{eq:eig_orthonormal_cols}
\\
\sum_{k = 1}^n \frac{ \phi_{ik} \, \phi_{jk} }{ p_k }
&= \delta_{ij}.
\label{eq:eig_orthonormal_rows}
\end{align}

@@

In analogous to Eq. \eqref{eq:x2t_average},
we have, from Eq. \eqref{eq:y2_sum},
the error after a period of $T$,
%
\begin{equation}
  \Err
  =
  \Err_R + \Err_A,
  \label{eq:error_split}
\end{equation}
%
which is the sum of the residual error
%
\begin{equation}
  \Err_R
  =
  \sum_{k = 2}^n
    \left\langle y_k^2(0) \right\rangle \,
    e^{ - 2 \, \lambda_k  \, q(T) },
  \label{eq:error_res}
\end{equation}
%
from the decay of the initial error,
and the asymptotic error
\begin{equation}
  \Err_A
  =
  \int_0^T \int_0^T
  \sum_{k = 2}^n
    \dot u_k\bigl( q(t) \bigr) \,
    \dot u_k\bigl( q(t') \bigr) \,
    \kappa_k(t - t') \, dt \, dt',
  \label{eq:error_asym}
\end{equation}
%
where,
\begin{align*}
  u_k(q')
  &\equiv
  e^{\lambda_k \, [q' - q(T)]},
  \\
  \kappa_k(t - t')
  &\equiv
  \left\langle
    \theta_k(t) \, \theta_k(t')
  \right\rangle.
\end{align*}
%
\note{The extremal condition of $\Err_A$ is given by
%
\begin{align}
\sum_{k = 2}^n
\lambda_k \, u_k\bigl( q(t) \bigr)
\int_0^T
  \ddot u_k \bigl( q(t') \bigr) \,
  \kappa_k(t' - t) \, dt' = 0.
\label{eq:optimal_mbin}
\end{align}
%
If all $\lambda_k$, hence all $u_k$, are the same,
the above condition can be satisfied
by Eq. \eqref{eq:ddu_eq_0},
which is the single-bin case.
%
Unfortunately, the above expression
appears to be useless in a general setting.
}

@@

In the asymptotic regime
(i.e., if the simulation length, $T$, is sufficiently long),
we can approximate $\kappa_k(t)$ as
%
\begin{equation}
  \kappa_k(t) \approx \Gamma_k \, \delta(t)
  \label{eq:kappa_delta}
\end{equation}
%
where,
\begin{equation}
  \Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt,
  \notag
  %\label{eq:Gamma_integral}
\end{equation}
or in discrete time,
\begin{equation}
  \Gamma_k = \sum_{t = -\infty}^\infty \kappa_k(t).
  \label{eq:Gamma_sum}
\end{equation}
%
Then, Eq. \eqref{eq:error_asym} is simplified as
%
\begin{align}
  \Err_A
  &=
  \int_0^T
  \sum_{k = 2}^n
  \Gamma_k \, \dot u_k^2\bigl( q(t) \bigr) \, dt.
  \label{eq:error_asym1}
\end{align}
%
\note{The next order correction to Eq. \eqref{eq:kappa_delta}
would lead to a correction to $\Err$
that is about $\alpha_{\max}^2$ times as large,
where $\alpha_{\max} = \max \alpha(t)$.
%
This is equivalent to saying that the Fourier transform
  $$
  \tilde \kappa_k(\omega) = \Gamma_k
  $$
  is a roughly constant.
  %
  For the next order correction we would have
  $$
  \tilde \kappa_k(\omega) = \Gamma_k + \Gamma^{(2)}_k \omega^2 + \dots,
  $$
  which leads to a correction to $\Err$
  $$
  \Gamma^{(2)}_k
  \int_0^T \ddot u_k^2\bigl( q(t) \bigr) \, dt
  \propto
  \alpha^2 \, \Err.
  $$
  This correction is of order $\alpha^2$
  compared to the main contribution in Eq. \eqref{eq:error_asym1}.
  %
  Thus, we can assume Eq. \eqref{eq:kappa_delta},
  once $\alpha(t)$ has dropped below a certain level.
}
%
\note{
  The extremal property, Eq. \eqref{eq:optimal_mbin},
  now becomes
  %
  \begin{align}
    \sum_{k = 1}^n
      \Gamma_k \, \lambda_k \,
      u_k\bigl( q(t) \bigr) \,
      \ddot u_k\bigl( q(t) \bigr) = 0.
    \label{eq:optimal_mbin1}
  \end{align}
  %
  If we interpret $u_k$ as the position,
  and $\Gamma_k \, \lambda_k$ as the mass,
  then the right-hand side gives the virial.
  %
  Thus, the extremal condition demands
  the total virial to be zero.
  $$\,$$
}
In this way,
the underlying sampling process
affects the error only through the set of numbers, $\Gamma_i$'s.

@@

This above matrix, $\mathbf w$,
being symmetric to the indices $i$ and $j$,
satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
for the flat distribution $p_i = 1/n$.
%
To find the eigenvectors,
we note that for a periodic variable $\varphi_i$,
the out-of-boundary values are defined by
%
\begin{equation}
  \varphi_i = \varphi_{i \pm n},
\label{eq:phi_pbc}
\end{equation}
%
such that $i \pm n$ lies in between $1$ and $n$.
%
We then have
%
\begin{equation}
  \sum_{ j = 1 }^n
    w_{ij} \, \varphi_j
  =
  \sum_{ j = 1 - n }^{ 2 \, n }
    \mu_{i - j} \, \varphi_j
  =
  \sum_{ l = -b }^{ b }
    \mu_l \, \varphi_{ i - l}
  .
\label{eq:wmul_to_convol}
\end{equation}
%
\note{The derivation of the first step of
  Eq. \eqref{eq:wmul_to_convol}:
$$
\begin{aligned}
  \sum_{j = 1}^n
    w_{ij} \, \varphi_j
  &=
  \sum_{j = 1}^n
    \mu_{i - j} \, \varphi_j
  +
  \sum_{j = 1}^n
    \mu_{i - j - n} \, \varphi_j
  +
  \sum_{j = 1}^n
    \mu_{i - j + n} \, \varphi_j
  \\
  &=
  \sum_{j = 1}^n
    \mu_{i - j} \, \varphi_j
  +
  \sum_{j = 1+n}^{2 \, n}
    \mu_{i - l} \, \varphi_l
  +
  \sum_{j = 1-n}^0
    \mu_{i - j + n} \, \varphi_j
  \\
  &=
  \sum_{j = 1-n}^{2 \, n}
    \mu_{i - j} \, \varphi_j
  .
\end{aligned}
$$
The second step of Eq. \eqref{eq:wmul_to_convol}
follows from the constraint $\mu_l = 0$ for $|l| > b$.

Similarly,
for the left eigenvector, we have
$$
  \sum_{ i = 1 }^n
    \varphi_i \, w_{ij}
  =
  \sum_{ l = -b }^b
    \varphi_{j - l} \, \mu_{-l}
  .
$$
But due to the symmetry Eq. \eqref{eq:mu_symm},
it is identical to \eqref{eq:wmul_to_convol}.
}

Thus,
we may construct a set of orthonormal eigenvectors,
$\pmb\varphi^{(1)}, \dots, \pmb\varphi^{(n)}$,
as
\begin{equation}
  \phi^{(k)}_j
  =
  \phi_{kj}
  =
  \frac{1}{\sqrt n}
  \exp
  \left[
    \ii
    \frac{ (k - 1) \, j \, 2 \, \pi }
         {            n             }
  \right]
  ,
  \notag
\end{equation}
%
with the eigenvalues being
%
\begin{equation}
  \lambda_k
  =
  \mu_0
  +
  2 \,
  \sum_{ l = 1 }^b
  \mu_l
  \cos
  \frac{ (k - 1) \, l \, 2 \, \pi }
       {            n             }
  .
  \label{eq:wband_eigenvalue_pbc}
\end{equation}
%
\note{Consider the unnormalized version
  $$
  \Phi^{(k)}_j =
  \exp\left[
    \frac{ ( k - 1 ) \, j \, 2 \, \pi }
         {              n             }
    \ii
  \right]
  ,
  $$
  we have
  $$
  \begin{aligned}
  \sum_{j = 1}^n
    w_{ij} \, \Phi^{(k)}_j
  &=
  \sum_{l = -b}^b
    \mu_l \, \Phi^{(k)}_{i - l}
  \\
  &=
  \mu_0 \, \Phi^{(k)}_i
  +
  \sum_{l = 1}^b
    \mu_l \,
    \left[ \Phi^{(k)}_{i - l} + \Phi^{(k)}_{i + l} \right]
  \\
  &=
  \Phi^{(k)}_i \,
  \left[
    \mu_0
    +
    2 \sum_{l = 1}^b
      \mu_l \, \cos
      \frac{ (k - 1) \, l \, 2 \, \pi }
           {            n             }
  \right]
  .
  \end{aligned}
  $$
  The term in the square brackets is the eigenvalue given by
  Eq. \eqref{eq:wband_eigenvalue_pbc}.
}
%
Note, however, the two-fold degeneracy as
$\lambda_{n - k + 2} = \lambda_k$.

@@

Similar to Eq. \eqref{eq:phi_pbc},
we define the out-of-boundary values
of $\varphi_j$ as
%
\begin{equation}
  \varphi_j
  =
  \begin{dcases}
    \varphi_{ 1 - j }           & \mathrm{for \;} j \le 0, \\
    \varphi_{ 2 \, n + 1 - j }  & \mathrm{for \;} j > n,
  \end{dcases}
\label{eq:phi_refl}
\end{equation}
%
such that Eq. \eqref{eq:wmul_to_convol}
still holds.
%
\note{The derivation of Eq. \eqref{eq:wmul_to_convol}
  in this case is similar,
  $$
  \begin{aligned}
    \sum_{j = 1}^n w_{ij} \, \varphi_j
    &=
    \sum_{j = 1}^n
      \mu_{i - j} \, \varphi_j
    +
    \sum_{j = 1}^n
      \mu_{i + j - 1} \, \varphi_j
    +
    \sum_{j = 1}^n
      \mu_{i + j - 2 \, n - 1} \, \varphi_j
    \\
    &=
    \sum_{j = 1}^n
      \mu_{i - j} \, \varphi_j
    +
    \sum_{l = 1 - n}^0
      \mu_{i - l} \, \varphi_l
    +
    \sum_{l = n + 1}^{ 2 \, n }
      \mu_{i - l} \, \varphi_l
    \\
    &=
    \sum_{j = 1 - n}^{ 2 \, n}
      \mu_{i - j} \, \varphi_j.
  \end{aligned}
  $$
}
%
Thus, we also have $n$ orthonormal eigenvectors,
$\pmb\varphi^{(1)}, \dots, \pmb\varphi^{(n)}$,
as
%
\begin{equation}
\varphi^{(k)}_i
=
\varphi_{k i}
=
\sqrt{
    \frac{ 2 - \delta_{k, 1} }
         {       n           }
     }
   \cos \frac{ ( k - 1) \, \left( i - \frac 1 2 \right) \, \pi}
            {                  n                         }.
\notag
%\label{eq:wband_eigenvector_refl}
\end{equation}
%
with eigenvalues
%
\begin{align}
  \lambda_k
  &=
  \mu_0
  +
  2
  \sum_{l = 1}^b
    \mu_l
    \cos \frac{(k - 1)  \, l \pi}{n}
  \label{eq:wband_eigenvalue_refl}
\end{align}
%
\note{We derive the above two equations below.

@@

  Now, for the orthogonality,
  $$
  \begin{aligned}
    \sum_{k = 1}^n
    \Phi^{(k)}_i \, \Phi^{(k)}_j
    &=
    \frac 1 2
    \sum_{k = 1}^n
    \left[
      \cos \tfrac{ (i - j) (k - 1) \, \pi }
                 {         n              }
      +
      \cos \tfrac{ (i + j - 1) (k - 1) \, \pi }
                 {             n              }
    \right]
    \\
    &=
    \frac 1 2
    \left[
      n \, \delta_{i, j}
      +
      \operatorname{odd}(i - j)
      +
      \operatorname{odd}(i + j - 1)
    \right]
    \\
    &=
    \frac n 2 \, \delta_{i, j}
    + \frac 1 2.
  \end{aligned}
  $$
  where we have used the fact
  that $i - j$ shares the parity with $i + j$
  in the last step.
  %
  So
  $$
    \sum_{k = 1}^n
    \varphi^{(k)}_i \, \varphi^{(k)}_j
    =
    \frac 2 n
    \sum_{k = 1}^n
    \Phi^{(k)}_i \, \Phi^{(k)}_j
    -
    \frac 1 n
    =
    \delta_{i, j}.
  $$
}
%

@@

In the above, we have been focusing on
the asymptotic error.
%
For a finite-length simulation,
we should include the residual part,
Eq. \eqref{eq:error_res}
and minimize the total error.
%
For the bandpass updating schemes,
this allows to determine the optimal value
of the free parameter, $t_0$.

Assuming the system is initially equilibrated
under a constant updating magnitude,
$\alpha_0$,
we may use Eq. \eqref{eq:error_res1} with
$$
q(T)
=
\int \frac{ dt } { t + t_0 }
=
\ln \frac{ T + t_0 } { t_0 }.
$$
%
Thus, from Eqs. \eqref{eq:error_split}
and \eqref{eq:error_asym_sinc}, we have
%
\begin{equation}
  E
  =
  E_R + E_A
  =
  \frac { \frac 1 2 \alpha_0 \, t_0^2 + T }
        {          ( T + t_0 )^2          }
  \sum_{ k = 2 }^K \Gamma_k
  .
\notag
%\label{eq:error_sinc0}
\end{equation}
%
The optimal error is reached at
%
\begin{equation}
  t_0
  =
  \frac{     2    }
       { \alpha_0 }
  ,
\label{eq:t0_sinc}
\end{equation}
%
and at this point
%
\begin{equation}
  E
  =
  \frac{        1         }
       { T + 2 / \alpha_0 }
  \sum_{ k = 2 }^K
    \Gamma_k
  .
\label{eq:error_sinc}
\end{equation}
%
This means that at the beginning of the schedule,
the updating magnitude is
$\alpha(0^+) = \alpha_0 / 2$,
half of the previous equilibrium value.
%
Interestingly, this factor of $1/2$
happens to coincide with the suggested
reduction factor of updating magnitude
in the original WL algorithm\cite{wang2001, wang2001pre}
at stage transitions.

>>>>

Going back to the problem of
minimizing Eq. \eqref{eq:error_sum}
for the single-bin scheme.
%
Note that Eq. \eqref{eq:dxdt_WL}
is of the form of Eq. \eqref{eq:dxdt_alpha}.
%
Thus, with $\lambda = 1$,
Eq. \eqref{eq:alpha_opt}
will optimize each term $\langle x_i^2 \rangle$,
and hence the sum.
%
In other words, the optimal $\alpha(t)$
is given by
\begin{equation}
\alpha(t)
=
\frac{    1     }
     {  t + t_0 }
,
\label{eq:alpha_invt1}
\end{equation}
and Eq. \eqref{eq:alpha_invt}
is the case of $t_0 = 0$.
%
In Sec. \ref{sec:finlencorr},
we show that $t_0$ can be adjusted
to take into account for the initial error.

@@

We now consider the multiple-bin scheme,
Eq. \eqref{eq:mbin_update}.
%
In this case,
upon a visit to bin $j$,
we update not only the bias potential at bin $j$,
but also those at a few neighboring bins, $i$'s.
%
The relative update magnitudes, $w_{ij}$,
constitute a matrix $\mathbf w$.
%
Then, Eq. \eqref{eq:vt_diffeq} is generalized to
\begin{equation}
  \dot v_i(t)
  =
  \alpha(t) \sum_{j=1}^n \, w_{ij} \frac{ h_j(t) } { p_j }.
  \label{eq:vt_diffeq_mbin}
\end{equation}

@@

The matrix $\mathbf w$ is not arbitrary.
%
A necessary condition to sample
the desired distribution
$\mathbf p = (p_1, \dots, p_n)$
is the following.
%
In Eq. \eqref{eq:vt_diffeq_mbin},
if $h_j(t)$ were the same as $p_j$,
the rates of change $\dot v_i(t)$'s
should be independent of $i$.
%
This allows $v_i(t)$'s to grow uniformly
in the asymptotic regime.
%
Thus, by a proper scaling of $\alpha(t)$,
we may write this condition as
%
\begin{equation}
  \sum_{j = 1}^n w_{ij} = 1.
  \label{eq:w_sumj}
\end{equation}
%
The above condition means that $(1, \dots, 1)^T$
is a right eigenvector of $\mathbf w$
with eigenvalue $1$.
%
Thus, the transpose $\mathbf w^T$
resembles a transition matrix,
although the elements can be negative.

@@

Further, by Eq. \eqref{eq:w_detailedbalance},
we can diagonalize $\mathbf w$ with a set of
eigenvectors
%
\begin{equation}
  \sum_{i = 1}^n \phi_{ki} \, w_{ij} = \lambda_k \, \phi_{kj},
  \label{eq:eig_w}
\end{equation}
%
satisfying the orthonormal conditions\cite{vankampen}:
%
\begin{align}
\sum_{k = 1}^n \phi_{ki} \, \phi_{kj}
&= \delta_{ij} \, p_i,
\label{eq:eig_orthonormal_cols}
\\
\sum_{k = 1}^n \frac{ \phi_{ik} \, \phi_{jk} }{ p_k }
&= \delta_{ij}.
\label{eq:eig_orthonormal_rows}
\end{align}
%
\note{We can define a symmetric matrix $\hat{\mathbf w}$
as $\hat w_{ij} = \sqrt{p_i/p_j} \, w_{ij}$,
which can be diagonalized\cite{vankampen}
with a set of orthonormal eigenvectors:
%
$\sum_{i = 1}^n \varphi_{ki} \, \hat w_{ij} = \lambda_k \, \varphi_{kj}$.
%
Then,
in terms of diagonalizing $\mathbf w$, we have
$$\phi_{ij} \equiv \sqrt{p_j} \, \varphi_{ij}$$
satisfying Eqs. \eqref{eq:eig_orthonormal_cols}
and \eqref{eq:eig_orthonormal_rows}.
%
}%

@@

In analogous to Eq. \eqref{eq:x2t_average},
we have, from Eq. \eqref{eq:y2_sum},
the error after a period of $T$,
%
\begin{equation}
  \Err
  =
  \Err_R + \Err_A,
  \label{eq:error_split}
\end{equation}
%
which is the sum of the residual error
%
\begin{equation}
  \Err_R
  =
  \sum_{k = 2}^n
    \left\langle y_k^2(0) \right\rangle \,
    e^{ - 2 \, \lambda_k  \, q(T) },
  \label{eq:error_res}
\end{equation}
%
from the decay of the initial error,
and the asymptotic error
\begin{equation}
  \Err_A
  =
  \int_0^T \int_0^T
  \sum_{k = 2}^n
    \dot u_k\bigl( q(t) \bigr) \,
    \dot u_k\bigl( q(t') \bigr) \,
    \kappa_k(t - t') \, dt \, dt',
  \label{eq:error_asym}
\end{equation}
%
where,
\begin{align*}
  u_k(q')
  &\equiv
  e^{\lambda_k \, [q' - q(T)]},
  \\
  \kappa_k(t - t')
  &\equiv
  \left\langle
    \theta_k(t) \, \theta_k(t')
  \right\rangle.
\end{align*}
%
\note{The extremal condition of $\Err_A$ is given by
%
\begin{align}
  \sum_{k = 2}^n
  \lambda_k \, u_k\bigl( q(t) \bigr)
  \int_0^T
    \ddot u_k \bigl( q(t') \bigr) \,
    \kappa_k(t' - t) \, dt' = 0.
  \tag{N1}
\label{eq:optimal_mbin}
\end{align}
%
If all $\lambda_k$, hence all $u_k$, are the same,
the above condition can be satisfied with $\ddot u_k = 0$,
which recovers the single-bin case.
%
Unfortunately, the above expression
appears to be useless in a general setting.
}

@@

In the asymptotic regime
(i.e., if the simulation length, $T$, is sufficiently long),
we can approximate $\kappa_k(t)$ as
%
\begin{equation}
  \kappa_k(t) \approx \Gamma_k \, \delta(t)
  \label{eq:kappa_delta}
\end{equation}
%
where,
\begin{equation}
  \Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt,
  \notag
  %\label{eq:Gamma_integral}
\end{equation}
or in discrete time,
\begin{equation}
  \Gamma_k = \sum_{t = -\infty}^\infty \kappa_k(t).
  \label{eq:Gamma_sum}
\end{equation}
%
Then, Eq. \eqref{eq:error_asym} is simplified as
%
\begin{align}
  \Err_A
  &=
  \int_0^T
  \sum_{k = 2}^n
  \Gamma_k \, \dot u_k^2\bigl( q(t) \bigr) \, dt.
  \label{eq:error_asym1}
\end{align}
%
\note{The next order correction to Eq. \eqref{eq:kappa_delta}
would lead to a correction to $\Err$
that is about $\alpha_{\max}^2$ times as large,
where $\alpha_{\max} = \max \alpha(t)$.
%
This is equivalent to saying that the Fourier transform
  $$
  \tilde \kappa_k(\omega) = \Gamma_k
  $$
  is a roughly constant.
  %
  For the next order correction we would have
  $$
  \tilde \kappa_k(\omega) = \Gamma_k + \Gamma^{(2)}_k \omega^2 + \dots,
  $$
  which leads to a correction to $\Err$
  $$
  \Gamma^{(2)}_k
  \int_0^T \ddot u_k^2\bigl( q(t) \bigr) \, dt
  \propto
  \alpha^2 \, \Err.
  $$
  This correction is of order $\alpha^2$
  compared to the main contribution in Eq. \eqref{eq:error_asym1}.
  %
  Thus, we can assume Eq. \eqref{eq:kappa_delta},
  once $\alpha(t)$ has dropped below a certain level.
}
%
\note{
  The extremal property, Eq. \eqref{eq:optimal_mbin},
  now becomes
  %
  \begin{align}
    \sum_{k = 1}^n
      \Gamma_k \, \lambda_k \,
      u_k\bigl( q(t) \bigr) \,
      \ddot u_k\bigl( q(t) \bigr) = 0
    .
    \tag{N2}
    \label{eq:optimal_mbin1}
  \end{align}
  %
  If we interpret $u_k$ as the position,
  and $\Gamma_k \, \lambda_k$ as the mass,
  then the right-hand side gives the virial.
  %
  Thus, the extremal condition demands
  the total virial to be zero.
  $$\,$$
}
In this way,
the underlying sampling process
affects the error only through the set of numbers, $\Gamma_i$'s.

@@

This above matrix, $\mathbf w$,
being symmetric to the indices $i$ and $j$,
satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
for the flat distribution $p_i = 1/n$.
%
To find the eigenvectors,
we note that for a periodic variable, $\phi_i$,
the out-of-boundary values are defined by
%
\begin{equation}
  \phi_i = \phi_{i \pm n},
\label{eq:phi_pbc}
\end{equation}
%
such that $i \pm n$ lies in between $1$ and $n$.
%
We then have
%
\begin{equation}
  \sum_{ j = 1 }^n
    w_{ij} \, \phi_j
  =
  \sum_{ j = 1 - n }^{ 2 \, n }
    \mu_{i - j} \, \phi_j
  =
  \sum_{ l = -b }^{ b }
    \mu_l \, \phi_{ i - l}
  .
\label{eq:wmul_to_convol}
\end{equation}
%
\note{The derivation of the first step of
  Eq. \eqref{eq:wmul_to_convol}:
$$
\begin{aligned}
  \sum_{j = 1}^n
    w_{ij} \, \phi_j
  &=
  \sum_{j = 1}^n
    \mu_{i - j} \, \phi_j
  +
  \sum_{j = 1}^n
    \mu_{i - j - n} \, \phi_j
  +
  \sum_{j = 1}^n
    \mu_{i - j + n} \, \phi_j
  \\
  &=
  \sum_{j = 1}^n
    \mu_{i - j} \, \phi_j
  +
  \sum_{j = 1+n}^{2 \, n}
    \mu_{i - l} \, \phi_l
  +
  \sum_{j = 1-n}^0
    \mu_{i - j + n} \, \phi_j
  \\
  &=
  \sum_{j = 1-n}^{2 \, n}
    \mu_{i - j} \, \phi_j
  .
\end{aligned}
$$
The second step of Eq. \eqref{eq:wmul_to_convol}
follows from the constraint $\mu_l = 0$ for $|l| > b$.

Similarly,
for the left eigenvector, we have
$$
  \sum_{ i = 1 }^n
    \phi_i \, w_{ij}
  =
  \sum_{ l = -b }^b
    \phi_{j - l} \, \mu_{-l}
  .
$$
But due to the symmetry Eq. \eqref{eq:mu_symm},
it is identical to \eqref{eq:wmul_to_convol}.
}

Thus,
we may construct a set of orthonormal eigenvectors,
$\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
as
\begin{equation}
  \phi^{(k)}_j
  =
  \phi_{kj}
  =
  \frac 1 n
  \exp
  \left[
    \ii
    \frac{ (k - 1) \, j \, 2 \, \pi }
         {            n             }
  \right]
  ,
  \notag
\end{equation}
%
with the eigenvalues being
%
\begin{equation}
  \lambda_k
  =
  \mu_0
  +
  2 \,
  \sum_{ l = 1 }^b
  \mu_l
  \cos
  \frac{ (k - 1) \, l \, 2 \, \pi }
       {            n             }
  .
  \label{eq:wband_eigenvalue_pbc}
\end{equation}
%
\note{Consider the unnormalized version
  $$
  \Phi^{(k)}_j =
  \exp\left[
    \frac{ ( k - 1 ) \, j \, 2 \, \pi }
         {              n             }
    \ii
  \right]
  ,
  $$
  we have
  $$
  \begin{aligned}
  \sum_{j = 1}^n
    w_{ij} \, \Phi^{(k)}_j
  &=
  \sum_{l = -b}^b
    \mu_l \, \Phi^{(k)}_{i - l}
  \\
  &=
  \mu_0 \, \Phi^{(k)}_i
  +
  \sum_{l = 1}^b
    \mu_l \,
    \left[ \Phi^{(k)}_{i - l} + \Phi^{(k)}_{i + l} \right]
  \\
  &=
  \Phi^{(k)}_i \,
  \left[
    \mu_0
    +
    2 \sum_{l = 1}^b
      \mu_l \, \cos
      \frac{ (k - 1) \, l \, 2 \, \pi }
           {            n             }
  \right]
  .
  \end{aligned}
  $$
  The term in the square brackets is the eigenvalue given by
  Eq. \eqref{eq:wband_eigenvalue_pbc}.
}
%
Note, however, the two-fold degeneracy as
$\lambda_{n - k + 2} = \lambda_k$.

@@

Similar to Eq. \eqref{eq:phi_pbc},
we define the out-of-boundary values
of $\phi_j$ as
%
\begin{equation}
  \phi_j
  =
  \begin{dcases}
    \phi_{ 1 - j }           & \mathrm{for \;} j \le 0, \\
    \phi_{ 2 \, n + 1 - j }  & \mathrm{for \;} j > n,
  \end{dcases}
\label{eq:phi_refl}
\end{equation}
%
such that Eq. \eqref{eq:wmul_to_convol}
still holds.
%
\note{The derivation of Eq. \eqref{eq:wmul_to_convol}
  in this case is similar,
  $$
  \begin{aligned}
    \sum_{j = 1}^n w_{ij} \, \phi_j
    &=
    \sum_{j = 1}^n
      \mu_{i - j} \, \phi_j
    +
    \sum_{j = 1}^n
      \mu_{i + j - 1} \, \phi_j
    +
    \sum_{j = 1}^n
      \mu_{i + j - 2 \, n - 1} \, \phi_j
    \\
    &=
    \sum_{j = 1}^n
      \mu_{i - j} \, \phi_j
    +
    \sum_{l = 1 - n}^0
      \mu_{i - l} \, \phi_l
    +
    \sum_{l = n + 1}^{ 2 \, n }
      \mu_{i - l} \, \phi_l
    \\
    &=
    \sum_{j = 1 - n}^{ 2 \, n}
      \mu_{i - j} \, \phi_j.
  \end{aligned}
  $$
}
%
Thus, we also have $n$ orthonormal eigenvectors,
$\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
as
%
\begin{equation}
  \phi^{(k)}_i
  =
  \phi_{k i}
  =
  \frac{ \sqrt{ 2 - \delta_{k, 1} } }
       {             n              }
  \cos \frac{ ( k - 1) \, \left( i - \frac 1 2 \right) \, \pi}
            {                    n                           }
  ,
\notag
%\label{eq:wband_eigenvector_refl}
\end{equation}
%
with eigenvalues
%
\begin{align}
  \lambda_k
  &=
  \mu_0
  +
  2
  \sum_{l = 1}^b
    \mu_l
    \cos \frac{(k - 1)  \, l \pi}{n}
  .
  \label{eq:wband_eigenvalue_refl}
\end{align}
%
\note{We derive the above two equations below.

@@

  Now, for the orthogonality,
  $$
  \begin{aligned}
    \sum_{k = 1}^n
    \Phi^{(k)}_i \, \Phi^{(k)}_j
    &=
    \frac 1 2
    \sum_{k = 1}^n
    \left[
      \cos \tfrac{ (i - j) (k - 1) \, \pi }
                 {         n              }
      +
      \cos \tfrac{ (i + j - 1) (k - 1) \, \pi }
                 {             n              }
    \right]
    \\
    &=
    \frac 1 2
    \left[
      n \, \delta_{i, j}
      +
      \operatorname{odd}(i - j)
      +
      \operatorname{odd}(i + j - 1)
    \right]
    \\
    &=
    \frac n 2 \, \delta_{i, j}
    + \frac 1 2.
  \end{aligned}
  $$
  where we have used the fact
  that $i - j$ shares the parity with $i + j$
  in the last step.
  %
  So
  $$
    \sum_{k = 1}^n
    \phi^{(k)}_i \, \phi^{(k)}_j
    =
    \frac 2 { n^2 }
    \sum_{k = 1}^n
    \Phi^{(k)}_i \, \Phi^{(k)}_j
    -
    \frac 1 { n^2 }
    =
    \frac 1 n \,
    \delta_{i, j}.
  $$
}
%

@@

In the above, we have been focusing on
the asymptotic error.
%
For a finite-length simulation,
we should include the residual part,
Eq. \eqref{eq:error_res}
and minimize the total error.
%
For the bandpass updating schemes,
this allows to determine the optimal value
of the free parameter, $t_0$.
%
%Assuming the system is initially equilibrated
%under a constant updating magnitude,
%$\alpha_0$,
Assuming Eq. \eqref{eq:error_res1} with
$$
q(T)
=
\int_0^T \frac{ dt } { t + t_0 }
=
\ln \frac{ T + t_0 } { t_0 }.
$$
%
we have, from Eqs. \eqref{eq:error_split}
and \eqref{eq:error_asym_sinc},
%
\begin{equation}
  E
  =
  E_R + E_A
  =
  \frac { \frac 1 2 \alpha_0 \, t_0^2 + T }
        {          ( T + t_0 )^2          }
  \sum_{ k = 2 }^K \Gamma_k
  .
\notag
%\label{eq:error_sinc0}
\end{equation}
%
The optimal error is reached at
%
\begin{equation}
  t_0
  =
  \frac{     2    }
       { \alpha_0 }
  ,
\label{eq:t0_sinc}
\end{equation}
%
and at this point
%
\begin{equation}
  E
  =
  \frac{        1         }
       { T + 2 / \alpha_0 }
  \sum_{ k = 2 }^K
    \Gamma_k
  .
\label{eq:error_sinc}
\end{equation}
%
This means that at the beginning of the schedule,
the updating magnitude is
$\alpha(0^+) = \alpha_0 / 2$,
half of the previous equilibrium value.
%
Interestingly, this factor of $1/2$
happens to coincide with the suggested
reduction factor of updating magnitude
in the original WL algorithm\cite{wang2001, wang2001pre}
at stage transitions.
